---
title: "Examples of using RCzechia"
author: "Jindra Lacko"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Solving real world issues with RCzechia}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```
## Geocoding locations & drawing them on a map

Drawing a map: three semi-random landmarks on map, with rivers and major water bodies shown for better orientation.

Geocoding using `geocode()` function from `ggmap` package, which implies consent with Google terms of use.

```{r geocode, options}
library(tidyverse)
library(RCzechia)
library(ggmap)
library(tmap)
library(sf)

reky <- reky[reky$Major == T, ]
plochy <- plochy[plochy$Major == T, ]


mista <- c('kramarova vila', 'arcibiskupske zahrady v kromerizi', 'becov nad teplou') 
  # locations to geocode

POI <- geocode(mista, output = "latlon") %>% # geocode using ggmap
  st_as_sf(coords = c("lon", "lat"), crs = 4326) # convert plain data to spatial

tm_plot <- tm_shape(republika) + tm_borders("grey30", lwd = 1) +
  tm_shape(POI) + tm_symbols(col = "firebrick3", shape = 20, size = 0.25) +
  tm_shape(reky) + tm_lines(col = 'steelblue', lwd = 1.5)  +
  tm_shape(plochy) + tm_polygons(fill = 'steelblue')
  

print(tm_plot)

```


## Unemployment in Czech Republic - a chloropleth

A visualization problem: unemployment in the Czech Republic is in general low, but not uniformly low.  
What are the hotspots?

```{r unempl, echo = T, eval = T, message = F}
library(tidyverse)
library(RCzechia)
library(tmap)
library(sf)

src <- read.csv(url("https://raw.githubusercontent.com/jlacko/RCzechia/master/data-raw/unempl.csv"), stringsAsFactors = F) 
# open data on unemployment from Czech Statistical Office - https://www.czso.cz/csu/czso/otevrena_data
# lightly edited for size (rows filtered)

src <- src %>%
  mutate(KOD_OBEC = as.character(uzemi_kod))  # keys in RCzechia are of type character

podklad <- obce_polygony %>% # obce_polygony = municipalities in RCzechia package
  inner_join(src, by = "KOD_OBEC") # linking by key


vystup <- tm_shape(republika) + tm_borders(col = "grey40") +
  tm_shape(podklad) + tm_fill(col = "hodnota", title = "Unemployment", palette = "YlOrRd") +
  tm_legend(position = c("RIGHT", "top"),
            legend.format = list(fun = function(x) paste0(formatC(x, digits = 0, format = "f"), " %")))

print(vystup)
 
```

## Distnance between Prague to Brno

Calculate distance between two spatial objects; the `sf` package supports (via gdal) point to point, point to polygon and polygon to polygon distances.  

Calculating distance from Prague (#1 Czech city) to Brno (#2 Czech city).

```{r distance, echo = T, eval = T, message = F}
library(tidyverse)
library(RCzechia)
library(sf)
library(units)

praha <- obce_polygony[obce_polygony$NAZ_OBEC == "Praha", ]
brno <- obce_polygony[obce_polygony$NAZ_OBEC == "Brno", ]

vzdalenost <- st_distance(praha, brno) %>%
  set_units("kilometers") # easier to interpret than meters, miles or decimal degrees..

print(vzdalenost)

```

## Interactive Map 

Interactive maps are powerful tools for data vizualization. They are easy to produce with the `tmap` package.

I found the stamen toner a good company for interactive chloropleths - it gives enough context without distracting from the story of your data.

A map of the whole Czech Republic in original resolution (the accuracy is about 1 meter) would be rather sizeable, and I found it better policy to either  
 * limit it to a single region (say a NUTS3) or  
 * to limit the size by applying `st_simplify()` to the shapefile. Note that RCzechia uses EPSG:4326 projection, with decimal degrees as unit. To simplify to given tolerance in meters you need to first  `st_transform()` it to a different projection, e.g. EPSG:5513 (ing. Křovák).


```{r interactive, echo = T, eval = T, message = F}
library(tidyverse)
library(RCzechia)
library(tmap)
library(sf)

src <- read.csv(url("https://raw.githubusercontent.com/jlacko/RCzechia/master/data-raw/unempl.csv"), stringsAsFactors = F) 
# open data on unemployment from Czech Statistical Office - https://www.czso.cz/csu/czso/otevrena_data
# lightly edited for size (rows filtered)


src <- src %>%
  mutate(KOD_OBEC = as.character(uzemi_kod))  # keys in RCzechia are of type character

podklad <- obce_polygony %>% # obce_polygony = municipalities in RCzechia package
  inner_join(src, by = "KOD_OBEC") %>% # linking by key
  filter(KOD_CZNUTS3 == "CZ071") # Olomoucký kraj

vystup <- tm_shape(republika) + tm_borders(col = "grey40") +
  tm_shape(podklad) + tm_fill(col = "hodnota", title = "Unemployment", palette = "YlOrRd", id = "NAZ_OBEC") +
  tm_legend(position = c("RIGHT", "top"),
            legend.format = list(fun = function(x) paste0(formatC(x, digits = 0, format = "f"), " %"))) +
  tm_view(basemaps = "Stamen.Toner")

save_tmap(vystup, filename = "vystup.html")
 
```


