---
title: "Solving real world problems with RCzechia"
author: "Jindra Lacko"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

## Unemployment in Czech Republic - a chloropleth

A visualization problem: unemployment in the Czech Republic is in general low, but not uniformly low.  

What are the hotspots?

```{r unempl, echo = T, eval = T, message = F}
library(tidyverse)
library(RCzechia)
library(tmap)
library(sf)

src <- read.csv(url("https://raw.githubusercontent.com/jlacko/RCzechia/sf-dev/data-raw/250169-17data011518.csv"), stringsAsFactors = F) # open data on unemployment from Czech Statistical Office - https://www.czso.cz/csu/czso/otevrena_data

src <- src %>%
  filter(rok == 2017 & mesic == 12 & vuk == "NEZ0004") %>% # December 2017 unemployment ratio only...
  mutate(KOD_OBEC = as.character(uzemi_kod))  # keys in RCzechia are of type character

podklad <- obce_polygony %>% # obce_polygony = municipalities in RCzechia package
  inner_join(src, by = "KOD_OBEC") # linking by key


tm_shape(republika) + tm_borders(col = "grey40") +
  tm_shape(podklad) + tm_fill(col = "hodnota", title = "Unemployment", palette = "YlOrRd") +
  tm_legend(position = c("RIGHT", "top"))
 
```

## Distnance - Prague to Brno

A common problem - calculate distance between two cities; the `sf` package supports point to point, point to polygon and polygon to polygon distances.  

Calculating distance from Prague (#1 Czech city) to Brno (#2 Czech city).

```{r distance, echo = T, eval = T, message = F}
library(tidyverse)
library(RCzechia)
library(sf)
library(units)

praha <- obce_polygony[obce_polygony$NAZ_OBEC == "Praha", ]
brno <- obce_polygony[obce_polygony$NAZ_OBEC == "Brno", ]

vzdalenost <- st_distance(praha, brno) %>%
  set_units("kilometers") # the default unit

print(vzdalenost)

```




