---
title: "Examples of using RCzechia"
author: "Jindra Lacko"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Solving real world issues with RCzechia}
  %\VignetteDepends{readxl}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
library(httr)
library(roxygen2)

```
### Czech Republic population

Population of the Czech Republic as per the [latest census](https://www.czso.cz/csu/czso/zakladni-vysledky-scitani-lidu-domu-a-bytu-2011-ceska-republika-2011-eihs9d4gp3) in 2011, per district (okres).


```{r census, echo = T, eval = T, message = F}
library(RCzechia)
library(dplyr)
library(readxl)
library(httr)
library(tmap)
library(sf)

GET("https://raw.githubusercontent.com/jlacko/RCzechia/master/data-raw/zvcr034.xls", 
    write_disk(tf <- tempfile(fileext = ".xls")))
      # original (ugly!!) format from the Czech Statistical Office

src <- read_excel(tf, sheet = 1, col_names = F) %>% # read in with fake column names
  select(NAZ_LAU1 = X__1, # key for the shapefile
         obyvatel = X__2) %>% # population (as text)
  mutate(obyvatel = suppressWarnings(as.double(obyvatel))) %>% 
    # convert from text to number; ignore errors
  mutate(NAZ_LAU1 = ifelse(NAZ_LAU1 == "Hlavní město Praha", "Praha", NAZ_LAU1)) 
    # rename Prague (from The Capital to a regular city)
  
okresni_data <- okresy("low") %>% # data shapefile
  inner_join(src, by = "NAZ_LAU1") 
    # key for data connection - note the use of inner (i.e. filtering) join

vystup <- tm_shape(okresni_data) + tm_fill(col = "obyvatel", title = "Population", 
                                           palette = "Blues", style = "quantile", n = 5) +
  tm_shape(okresni_data) + tm_borders("grey40", lwd = 0.5) + # thin edges of districts
  tm_shape(republika("low")) + tm_borders("grey30", lwd = 1.5) # thick national borders

print(vystup)

```

## Geocoding locations & drawing them on a map

Drawing a map: three semi-random landmarks on map, with rivers shown for better orientation.

To get the geocoded data frame you can consider using `geocode()` function from `ggmap` package, which implies consent with Google terms of use.

```{r geocode, echo = T, eval = T, message = F}
library(RCzechia)
library(dplyr)
library(tmap)
library(sf)

rivers <- reky()

rivers <- rivers[rivers$Major == T, ]

mista <- data.frame(misto = c('kramarova vila', 'arcibiskupske zahrady v kromerizi', 'becov nad teplou'),
                    lon = c(14.41030, 17.39353, 12.83833),
                    lat = c(50.09380, 49.30048, 50.08346))

# to geocode a list of locations consider ggmap::geocode()

POI <- mista %>% # or geocode using ggmap
  st_as_sf(coords = c("lon", "lat"), crs = 4326) # convert plain data to spatial CRS = WGS84, used by Google

tm_plot <- tm_shape(republika("low")) + tm_borders("grey30", lwd = 1) +
  tm_shape(POI) + tm_symbols(col = "firebrick3", shape = 20, size = 0.5) +
  tm_shape(rivers) + tm_lines(col = 'steelblue', lwd = 1.5)
  

print(tm_plot)

```


## Unemployment in Czech Republic - a chloropleth

A visualization problem: unemployment in the Czech Republic is in general low, but not uniformly low.  
What are the hotspots?

```{r unempl, echo = T, eval = T, message = F}
library(dplyr)
library(RCzechia)
library(tmap)
library(sf)

src <- read.csv(url("https://raw.githubusercontent.com/jlacko/RCzechia/master/data-raw/unempl.csv"), stringsAsFactors = F) 
# open data on unemployment from Czech Statistical Office - https://www.czso.cz/csu/czso/otevrena_data
# lightly edited for size (rows filtered)

src <- src %>%
  mutate(KOD_OBEC = as.character(uzemi_kod))  # keys in RCzechia are of type character

podklad <- obce_polygony() %>% # obce_polygony = municipalities in RCzechia package
  inner_join(src, by = "KOD_OBEC") # linking by key


vystup <- tm_shape(republika()) + tm_borders(col = "grey40") +
  tm_shape(podklad) + tm_fill(col = "hodnota", title = "Unemployment", palette = "YlOrRd") +
  tm_legend(position = c("RIGHT", "top"),
            legend.format = list(fun = function(x) paste0(formatC(x, digits = 0, format = "f"), " %")))

print(vystup)
 
```

## Distnance between Prague to Brno

Calculate distance between two spatial objects; the `sf` package supports (via gdal) point to point, point to polygon and polygon to polygon distances.  

Calculating distance from Prague (#1 Czech city) to Brno (#2 Czech city).

```{r distance, echo = T, eval = T, message = F}
library(dplyr)
library(RCzechia)
library(sf)
library(units)

obce <- obce_polygony()

praha <- obce[obce$NAZ_OBEC == "Praha", ]
brno <- obce[obce$NAZ_OBEC == "Brno", ]

vzdalenost <- st_distance(praha, brno) %>%
  set_units("kilometers") # easier to interpret than meters, miles or decimal degrees..

print(vzdalenost)

```

## Interactive Map 

Interactive maps are powerful tools for data vizualization. They are easy to produce with the `tmap` package.

I found the stamen toner a good company for interactive chloropleths - it gives enough context without distracting from the story of your data.

A map of the whole Czech Republic in original resolution (the accuracy is about 1 meter) would be rather sizeable, and I found it better policy to either  
 * limit it to a single region (say a NUTS3) or  
 * to limit the size by applying `st_simplify()` to the shapefile. Note that RCzechia uses EPSG:4326 projection, with decimal degrees as unit. To simplify to given tolerance in meters you need to first  `st_transform()` it to a different projection, e.g. EPSG:5513 (ing. Křovák).


```{r interactive, echo = T, eval = T, message = F}
library(dplyr)
library(RCzechia)
library(tmap)
library(sf)

src <- read.csv(url("https://raw.githubusercontent.com/jlacko/RCzechia/master/data-raw/unempl.csv"), stringsAsFactors = F) 
# open data on unemployment from Czech Statistical Office - https://www.czso.cz/csu/czso/otevrena_data
# lightly edited for size (rows filtered)


src <- src %>%
  mutate(KOD_OBEC = as.character(uzemi_kod))  # keys in RCzechia are of type character

podklad <- obce_polygony() %>% # obce_polygony = municipalities in RCzechia package
  inner_join(src, by = "KOD_OBEC") %>% # linking by key
  filter(KOD_CZNUTS3 == "CZ071") # Olomoucký kraj

vystup <- tm_shape(republika()) + tm_borders(col = "grey40") +
  tm_shape(podklad) + tm_fill(col = "hodnota", title = "Unemployment", palette = "YlOrRd", id = "NAZ_OBEC") +
  tm_legend(position = c("RIGHT", "top"),
            legend.format = list(fun = function(x) paste0(formatC(x, digits = 0, format = "f"), " %"))) +
  tm_view(basemaps = "Stamen.Toner")

save_tmap(vystup, filename = "vystup.html")
 
```

